---
title: "linear_modeling_moonlander_eye_data"
author: "Nils Wendel Heinrich"
date: '2023-02-17'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(ggplot2)
library(fitdistrplus)
```


```{r data, include=FALSE}
#current_wd = getwd()
data_fixations <- read_csv("eye_data/experimental_eye_data_fixations.csv")
data_saccades <- read_csv("eye_data/experimental_eye_data_saccades.csv")
```

```{r descriptives, include=FALSE}

unique(data_fixations$ID)
unique(data_saccades$ID)

```

```{r edit_data, include=TRUE}

## level difficulty

# fixations
data_fixations <- data_fixations %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
data_fixations$level_difficulty <- as.factor(data_fixations$level_difficulty)
data_fixations <- within(data_fixations, level_difficulty <- relevel(level_difficulty, ref = "easy"))

# saccades
data_saccades <- data_saccades %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
data_saccades$level_difficulty <- as.factor(data_saccades$level_difficulty)
data_saccades <- within(data_saccades, level_difficulty <- relevel(level_difficulty, ref = "easy"))


```

# linear modeling
```{r random_effects, include=TRUE}

# random intercept for ID

null_fixDur <- lmer(fixation_duration ~ 1 + (1|ID), data=data_fixations, REML=FALSE)
summary(null_fixDur)

null_fixLoc <- lmer(converging_eye_y ~ 1 + (1|ID), data=data_fixations, REML=FALSE)
summary(null_fixLoc)

null_saccAmp <- lmer(saccade_amplitude ~ 1 + (1|ID), data=data_saccades, REML=FALSE)
summary(null_saccAmp)

```
```{r ICCs, include=TRUE}

ICC_fixDur <- 0.02083 / (0.02083 + 0.08969)
ICC_fixDur

ICC_convergingEyeY <- 1927 / (1927 + 41220)
ICC_convergingEyeY

ICC_saccAmp <- 20822 / (20822 + 1349405)
ICC_saccAmp

```

## Fixations

The VPixx eye tracker might have detected fixations where there are none, resulting in a fixation duration of 0 ms. We will filter those and divide the data into fixations exploring the scene, and fixations that "rest" directly in front of the space ship.
```{r filtering, include=TRUE}

# only considering fixations with duration > 0:
data_fixations <- data_fixations[data_fixations$fixation_duration > 0, ]

# subsetting exploring fixations
explorative_fixations <- data_fixations[data_fixations$exploring_fixation == 1, ]
# vs. fixations resting in front of spaceship
resting_fixations <- data_fixations[data_fixations$exploring_fixation == 0, ]

```


```{r random_effects_for_explorative, include=TRUE}

# ID

null_1 <- lmer(fixation_duration ~ 1 + (1|ID), data=explorative_fixations, REML=FALSE)
summary(null_1)

null_2 <- lmer(converging_eye_y ~ 1 + (1|ID), data=explorative_fixations, REML=FALSE)
summary(null_2)


```

```{r ICCs_explore, include=TRUE}

ICC_fixDur_explore <- 0.01510 / (0.01510 + 0.08911)
ICC_fixDur_explore

ICC_convergingEyeY_explore <- 5501 / (5501 + 27074)
ICC_convergingEyeY_explore


```

# fixation duration

We will first take a look at fixation duration. From experience I know that duration are not normally distributed and naturally can not have negative values. We will explore how the data fits a gamma distribution.

```{r check_fixDur_fit, include=TRUE}

fit.gamma <- fitdist(data_fixations$fixation_duration, distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)
hist(data_fixations$fixation_duration, breaks=200)
```

The fit looks good. We will therefore use generalized linear modeling to fit a model on the gamma distributed fixation duration for exploring and resting fixations individually. We specify drift, input_noise, and level as fixed effects while assuming an interaction effect between drift and input_noise. 
```{r fixation_duration, include=TRUE}

# for explorative fixations
fixDur_explore <- glmer(fixation_duration ~ drift*level_difficulty + input_noise + (1|ID),  data = explorative_fixations, family = Gamma)
summary(fixDur_explore)

#fixDur_explore.free_slope <- glmer(fixation_duration ~ drift*level_difficulty + input_noise + (1 + input_noise|ID),  #data = explorative_fixations, family = Gamma)
#summary(fixDur_explore.free_slope)

#coef(fixDur_explore)
#coef(fixDur_explore.free_slope). # when freeing slope for input noise, we can actually see single individuals skewing the coefficients for input noise

# for resting fixations
fixDur_rest <- glmer(fixation_duration ~ drift*level_difficulty + input_noise + (1|ID),  data = resting_fixations, family = Gamma)
summary(fixDur_rest)

```

Discussing the results: ...

# fixation location

Next we will take a look at the fixation location in terms of y location (vertical position of eyes on screen). Fixation locations might actually be normally distributed. 
```{r check_fixloc, include=TRUE}
fit.gauss <- fitdist(data_fixations$converging_eye_y, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)
hist(data_fixations$converging_eye_y, breaks=200)
```

That looks smooth. We will therefore be able to fit a LMM without having to specify a family for our outcome variable converging_eye_y. 
We stick to our fixed effects model structure and fit a model for exploring and resting fixations individually.
```{r fixation_location, include=TRUE}

# subsetting exploring fixations
fixloc_explore <- lmer(converging_eye_y ~ drift*level_difficulty + input_noise + (1|ID),  data = explorative_fixations, REML = FALSE)
summary(fixloc_explore)

# vs. fixations resting in front of spaceship
fixloc_rest <- lmer(converging_eye_y ~ drift*level_difficulty + input_noise + (1|ID),  data = resting_fixations, REML = FALSE)
summary(fixloc_rest)

```

Discussing the results: ...

## Saccades

```{r filtering_saccades, include=TRUE}

# only considering fixations with duration > 0:
data_saccades <- data_saccades[data_saccades$saccade_amplitude > 0 & data_saccades$saccade_amplitude < 1000, ]
data_saccades <- na.omit(data_saccades)

# subsetting progressive saccades
progressive_saccades <- data_saccades[data_saccades$saccade_direction_y < 0, ]
# vs. regressive saccades going back to the spaceship
regressive_saccades <- data_saccades[data_saccades$saccade_direction_y > 0, ]

```

# saccade amplitude

```{r check_saccamp, include=TRUE}

fit.gamma <- fitdist(data_saccades$saccade_amplitude, distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

# sum(is.nan(data_saccades$saccade_amplitude))
hist(data_saccades$saccade_amplitude, breaks=200)
min(data_saccades$saccade_amplitude)
```

```{r saccade_amplitude, include=TRUE}

# subsetting exploring fixations
saccamp_progress <- lmer(saccade_amplitude ~ drift*level_difficulty + input_noise + (1|ID),  data = progressive_saccades, REML = FALSE)
summary(saccamp_progress)

# vs. fixations resting in front of spaceship
saccamp_regress <- lmer(saccade_amplitude ~ drift*level_difficulty + input_noise + (1|ID),  data = regressive_saccades, REML = FALSE)
summary(saccamp_regress)

```

Follow-up: Does saccades land nearer spaceship after saccade during strong input noise (compared to intercept)?

```{r soc, include=TRUE}

soc_data <- read_csv("aggregated_soc_data.csv")

soc_data <- soc_data %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
soc_data$level_difficulty <- as.factor(soc_data$level_difficulty)
soc_data <- within(soc_data, level_difficulty <- relevel(level_difficulty, ref = "easy"))

hist(soc_data$SoC, breaks=7)

```

```{r soc_model, include=TRUE}

# fixed effects structure to be discussed:
# 1) done + drift*level_difficulty + input_noise
# 2) drift*level_difficulty + input_noise + (1|done)

soc_model <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1|ID),  data = soc_data)
summary(soc_model)
coef(soc_model)
```