---
title: "linear_modeling_moonlander_eye_data"
author: "Nils Wendel Heinrich"
date: '2023-02-17'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(latex2exp)
library(ggplot2)
library(fitdistrplus)
library(simr)
```


```{r eye_data, include=FALSE}

setwd('/Users/heinrich/Projects/eye_movement_analysis_moonlander/')

data_fixations <- read_csv("eye_data/experimental_eye_data_fixations.csv")
data_saccades <- read_csv("eye_data/experimental_eye_data_saccades.csv")

```
We have data from 27 participants. Exclusion took place and will have to be specified here...

```{r edit_data, include=FALSE}

# level difficulty
# Most likely we won't use level difficulty as fixed effect. But we will nevertheless relevel it.
## fixations
data_fixations <- data_fixations %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
data_fixations$level_difficulty <- as.factor(data_fixations$level_difficulty)
data_fixations <- within(data_fixations, level_difficulty <- relevel(level_difficulty, ref = "easy"))
## saccades
data_saccades <- data_saccades %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
data_saccades$level_difficulty <- as.factor(data_saccades$level_difficulty)
data_saccades <- within(data_saccades, level_difficulty <- relevel(level_difficulty, ref = "easy"))

# input noise
## fixations
data_fixations$input_noise <- as.factor(data_fixations$input_noise)
data_fixations <- within(data_fixations, input_noise <- relevel(input_noise, ref = "N"))
## saccades
data_saccades$input_noise <- as.factor(data_saccades$input_noise)
data_saccades <- within(data_saccades, input_noise <- relevel(input_noise, ref = "N"))

```

```{r plotting_theme, echo=FALSE}

set_theme(
  base = theme_bw(),
  geom.outline.color = "antiquewhite4",
  legend.inside = TRUE,
  geom.label.size = 3
)

```

# Effects coding

We are especially interested in exlporing effects of *input noise* in our predicted variables. Input noise means the internal loss of control. We will put it against *drift*, the external loss of control as well as account for the number of obstacles on the screen which might mainly drive changes in the predicted variable.

We want to apply our own effects coding for the variables of level_difficulty and input_noise.
```{r effects_coding, include=TRUE}

# both variables have 3 levels. Therefore we can go for a single matrix. We want to test level 2 against level 1, then level 3 against level 1, and last level 3 against level 2.
#effects_coding <- contr.treatment(3)

#effects_coding <- cbind( c(0,1,0), c(0,0,1), c(0,-1,1))
#contrasts(data_fixations$input_noise) = effects_coding

# this changes the base to level 2
#data_fixations$input_noise <- as.factor(data_fixations$input_noise)
#contrasts(data_fixations$input_noise) = contr.treatment(3, base = 2)

```

# Predicting fixations

Fixations may have served various functions. We will try to account for it and differentiate between resting and exploring fixations. 
```{r splitting_fixations, include=TRUE}

# only considering fixations with duration > 0:
data_fixations <- data_fixations[data_fixations$fixation_duration > 0, ]
# The eye tracker might have flagged single frames as fixations. These would have lasted no time and we will simply filter them out.

# distinguishing exploring and resting fixations
#x_sp = 972
#y_sp = 288

# 4° is half of the span of the parafovea. We will use that as the threshold (keeping the spaceship clearly visible vs. not)
#threshold = 4

#data_fixations$distance_to_spaceship <- sqrt((data_fixations$converging_eye_x_adjusted - x_sp)**2 + (data_fixations$converging_eye_y_adjusted - y_sp)**2)

exploring_fixations <- data_fixations[data_fixations$exploring_fixation == 1, ]
resting_fixations <- data_fixations[data_fixations$exploring_fixation == 0, ]

```

## Resting fixations - Fixation duration (s)

```{r check_fixDur_rest_fit, include=TRUE}

hist(resting_fixations$fixation_duration, breaks=200)

fit.gamma <- fitdist(resting_fixations$fixation_duration, distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

```

```{r random_effects, include=TRUE}

fixDur_rest_null <- glmer(fixation_duration ~ 1 + (1|ID), data=resting_fixations, family = Gamma)
summary(fixDur_rest_null)

```

```{r ICC_fixDur_rest, include=TRUE}

ICC_fixDur_rest <- 1.653 / (1.653 + 19.262)
ICC_fixDur_rest

```

```{r fixation_duration_rest, include=TRUE}

pred_fixDur_rest <- glmer(fixation_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=resting_fixations, family = Gamma)

pred_fixDur_rest.slope <- glmer(fixation_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=resting_fixations, family = Gamma)

pred_fixDur_rest.slope2 <- glmer(fixation_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=resting_fixations, family = Gamma)

anova(pred_fixDur_rest, pred_fixDur_rest.slope, pred_fixDur_rest.slope2)

summary(pred_fixDur_rest.slope)

#powerSim(pred_fixDur_rest.slope, "input_noise", nsim=100)
#powerSim(pred_fixDur_rest.slope, nsim = 100, test = fixed("input_noise", method = "kr"))

```

```{r plot_fixation_duration_rest, echo=FALSE}

plot_fixDurRest <- plot_model(pred_fixDur_rest.slope, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_fixDurRest <- plot_fixDurRest + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Fixation Duration (s)")

plot_fixDurRest
```

## Resting fixations - Fixation location (distance to spaceship in visual degrees)

```{r check_distSpaceship_rest_fit, include=TRUE}

hist(resting_fixations$distance_to_spaceship, breaks=200)

fit.gauss <- fitdist(c(na.exclude(resting_fixations$distance_to_spaceship)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distSpaceship_rest_null <- lmer(distance_to_spaceship ~ 1 + (1|ID), data=resting_fixations, REML=FALSE)
summary(distSpaceship_rest_null)

```

```{r ICC_distSpaceship_rest, include=TRUE}

ICC_distSpaceship_rest <- 0.08396 / (0.08396 + 1.01570)
ICC_distSpaceship_rest

```

```{r distance_to_spaceship_rest, include=TRUE}

pred_distSpaceship_rest <- lmer(distance_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=resting_fixations, REML = FALSE)

pred_distSpaceship_rest.slope <- lmer(distance_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=resting_fixations, REML = FALSE)

pred_distSpaceship_rest.slope2 <- lmer(distance_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=resting_fixations, REML = FALSE)  # does not converge

anova(pred_distSpaceship_rest, pred_distSpaceship_rest.slope, pred_distSpaceship_rest.slope2)

summary(pred_distSpaceship_rest.slope)

```

```{r plot_distance_to_spaceship_rest, echo=FALSE}

plot_distSpaceshipRest <- plot_model(pred_distSpaceship_rest.slope, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_distSpaceshipRest <- plot_distSpaceshipRest + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Distance to Spaceship (visual degrees°)")

plot_distSpaceshipRest
```

Discussing the results: ...

## Exploring fixations - Fixation duration (s)

```{r check_fixDur_explore_fit, include=TRUE}

hist(exploring_fixations$fixation_duration, breaks=200)

fit.gamma <- fitdist(exploring_fixations$fixation_duration, distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

```

```{r random_effects, include=TRUE}

fixDur_explore_null <- glmer(fixation_duration ~ 1 + (1|ID), data=exploring_fixations, family = Gamma)
summary(fixDur_explore_null)

```

```{r ICC_fixDur_explore, include=TRUE}

ICC_fixDur_explore <- 2.039 / (2.039 + 13.707)
ICC_fixDur_explore

```

```{r fixation_duration_explore, include=TRUE}

pred_fixDur_explore <- glmer(fixation_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=exploring_fixations, family = Gamma)

pred_fixDur_explore.slope <- glmer(fixation_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=exploring_fixations, family = Gamma)

pred_fixDur_explore.slope2 <- glmer(fixation_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=exploring_fixations, family = Gamma)

pred_fixDur_explore.slope3 <- glmer(fixation_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=exploring_fixations, family = Gamma)

anova(pred_fixDur_explore, pred_fixDur_explore.slope, pred_fixDur_explore.slope2, pred_fixDur_explore.slope3)

summary(pred_fixDur_explore)

```

```{r plot_fixation_duration_explore, echo=FALSE}

plot_fixDur_explore <- plot_model(pred_fixDur_explore, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_fixDur_explore <- plot_fixDur_explore + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Fixation duration (s)")

plot_fixDur_explore
```

Discussing the results: ...

## Exploring fixations - Fixation location (distance to spaceship in visual degrees)

```{r check_distSpaceship_explore_fit, include=TRUE}

hist(exploring_fixations$distance_to_spaceship, breaks=200)

fit.gamma <- fitdist(c(na.exclude(exploring_fixations$distance_to_spaceship)), distr = "exp", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

#lnorm: AIC:  374016.8   BIC:  374034.8
#gamma: AIC:  400144.8   BIC:  400162.8
#exp: AIC:  407156.6   BIC:  407165.6



```

```{r random_effects, include=TRUE}

distSpaceship_explore_null <- glmer(distance_to_spaceship ~ 1 + (1|ID), data=exploring_fixations, family = Gamma)
summary(distSpaceship_explore_null)

```

```{r ICC_distSpaceship_explore, include=TRUE}

ICC_distSpaceship_explore <- 0.0003975 / (0.0003975 + 0.3856651)
ICC_distSpaceship_explore

```

```{r distance_to_spaceship_explore, include=TRUE}

# ID does not explain variance in the predicted variable. We might therefore omit including it as random effect
pred_distSpaceship_explore <- glm(distance_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise, data=exploring_fixations, family = Gamma)
summary(pred_distSpaceship_explore)

# including random effects
#pred_distSpaceship_explore <- lmer(distance_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=exploring_fixations, REML = FALSE)
#summary(pred_distSpaceship_explore)

#pred_distSpaceship_explore.slope <- lmer(distance_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=exploring_fixations, REML = FALSE)
#summary(pred_distSpaceship_explore.slope)

#anova(pred_distSpaceship_explore, pred_distSpaceship_explore.slope)

#pred_distSpaceship_explore.slope2 <- lmer(distance_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=exploring_fixations, REML = FALSE)
#summary(pred_distSpaceship_explore.slope2)

#anova(pred_distSpaceship_explore, pred_distSpaceship_explore.slope, pred_distSpaceship_explore.slope2)

#summary()

```

```{r plot_distance_to_spaceship_explore, echo=FALSE}

plot_distSpaceship_explore <- plot_model(pred_distSpaceship_explore, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_distSpaceship_explore <- plot_distSpaceship_explore + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Distance to Spaceship (visual degrees°)")

plot_distSpaceship_explore
```

Discussing the results: ...

# Predicting saccades

Saccades may have served various functions. We will try to account for it and differentiate between progressive and regressive saccades. 
```{r splitting_saccades, include=TRUE}

data_saccades <- data_saccades[data_saccades$saccade_amplitude > 0 & data_saccades$saccade_amplitude < 27, ]
# filter saccades that have no saccade amplitude
data_saccades <- data_saccades[!is.na(data_saccades$saccade_amplitude),]

progressive_saccades <- data_saccades[data_saccades$saccade_direction_y < 0, ]
regressive_saccades <- data_saccades[data_saccades$saccade_direction_y > 0, ]

```

## Progressive saccades - Saccade amplitude (in visual degrees)

```{r check_saccAmp_progress_fit, include=TRUE}

hist(progressive_saccades$saccade_amplitude, breaks=200)

fit.gamma <- fitdist(progressive_saccades$saccade_amplitude, distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

```

```{r random_effects, include=TRUE}

saccAmp_progress_null <- glmer(saccade_amplitude ~ 1 + (1|ID), data=progressive_saccades, family = Gamma)
summary(saccAmp_progress_null)

```

```{r ICC_saccAmp_progress, include=TRUE}

ICC_saccAmp_progress <- 0.008486 / (0.008486 + 1.225939)
ICC_saccAmp_progress

```

```{r saccAmp_progress, include=TRUE}

# ID does not explain variance in the predicted variable. We might therefore omit including it as random effect
pred_saccAmp_progress <- glm(saccade_amplitude ~ N_visible_obstacles * N_visible_drift_tiles + input_noise, data=progressive_saccades, family = Gamma)
summary(pred_saccAmp_progress)

```

```{r plot_saccAmp_progress, echo=FALSE}

plot_saccAmp_progress <- plot_model(pred_saccAmp_progress, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_saccAmp_progress <- plot_saccAmp_progress + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Saccade amplitude (visual degrees°)")

plot_saccAmp_progress
```

Discussing the results: ...

## Progressive saccades - distance to closest obstacle (in pixel)

```{r check_distObs_progress_fit, include=TRUE}

hist(progressive_saccades$saccLandSite_dist_to_closestObstacle, breaks=200)

fit.gamma <- fitdist(c(na.exclude(progressive_saccades$saccLandSite_dist_to_closestObstacle)), distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

```

```{r random_effects, include=TRUE}

distObs_explore_null <- glmer(saccLandSite_dist_to_closestObstacle ~ 1 + (1|ID), data=progressive_saccades, family = Gamma)
summary(distObs_explore_null)

```

```{r ICC_distObs_progress, include=TRUE}

ICC_distObs_progress <- 0.00183 / (0.00183 + 0.79070)
ICC_distObs_progress

```

```{r distObs_progress, include=TRUE}

# ID does not explain variance in the predicted variable. We might therefore omit including it as random effect
pred_distObs_progress <- glm(saccLandSite_dist_to_closestObstacle ~ N_visible_obstacles * N_visible_drift_tiles + input_noise, data=progressive_saccades, family = Gamma)
summary(pred_distObs_progress)

```

```{r plot_distance_to_obstacles_progress, echo=FALSE}

plot_distObs_progress <- plot_model(pred_distObs_progress, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_distObs_progress <- plot_distObs_progress + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Distance to nearest obstacle (in pixel)")

plot_distObs_progress
```

Discussing the results: ...

## Regressive saccades - Saccade amplitude

```{r check_saccAmp_regress_fit, include=TRUE}

hist(regressive_saccades$saccade_amplitude, breaks=200)

fit.gamma <- fitdist(regressive_saccades$saccade_amplitude, distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

```

```{r random_effects, include=TRUE}

saccAmp_regress_null <- glmer(saccade_amplitude ~ 1 + (1|ID), data=regressive_saccades, family = Gamma)
summary(saccAmp_regress_null)

```

```{r ICC_saccAmp_regress, include=TRUE}

ICC_saccAmp_regress <- 0.06734 / (0.06734 + 2.29774)
ICC_saccAmp_regress

```

```{r saccAmp_regress, include=TRUE}

# ID does not explain variance in the predicted variable. We might therefore omit including it as random effect
pred_saccAmp_regress <- glm(saccade_amplitude ~ N_visible_obstacles * N_visible_drift_tiles + input_noise, data=regressive_saccades, family = Gamma)
summary(pred_saccAmp_regress)

```

```{r plot_saccAmp_regress, echo=FALSE}

plot_saccAmp_regress <- plot_model(pred_saccAmp_regress, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_saccAmp_regress <- plot_saccAmp_regress + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Saccade amplitude (visual degrees°)")

plot_saccAmp_regress
```

Discussing the results: ...

## Regressive saccades - distance to spaceship

```{r check_distSpaceship_regress_fit, include=TRUE}

hist(regressive_saccades$saccLandSite_dist_to_spaceship, breaks=200)

fit.gauss <- fitdist(regressive_saccades$saccade_amplitude, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distSpaceship_regress_null <- lmer(saccLandSite_dist_to_spaceship ~ 1 + (1|ID), data=regressive_saccades, REML = FALSE)
summary(distSpaceship_regress_null)

```

```{r ICC_distSpaceship_regress, include=TRUE}

ICC_distSpaceship_regress <- 54.95 / (54.95 + 236.04)
ICC_distSpaceship_regress

```

```{r distSpaceship_regress, include=TRUE}

pred_distSpaceship_regress <- lmer(saccLandSite_dist_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=regressive_saccades, REML = FALSE)

pred_distSpaceship_regress.slope <- lmer(saccLandSite_dist_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=regressive_saccades, REML = FALSE) # does not converge

pred_distSpaceship_regress.slope2 <- lmer(saccLandSite_dist_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=regressive_saccades, REML = FALSE)

pred_distSpaceship_regress.slope3 <- lmer(saccLandSite_dist_to_spaceship ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=regressive_saccades, REML = FALSE)

anova(pred_distSpaceship_regress, pred_distSpaceship_regress.slope, pred_distSpaceship_regress.slope2, pred_distSpaceship_regress.slope3)

summary(pred_distSpaceship_regress.slope3)

```

```{r plot_distance_to_spaceship_regress, echo=FALSE}

plot_distSpaceship_regress <- plot_model(pred_distSpaceship_regress.slope3, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_distSpaceship_regress <- plot_distSpaceship_regress + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Distance to Spaceship (visual degrees°)")

plot_distSpaceship_regress
```

Discussing the results: ...

# Input data - input duration

```{r input_data, include=FALSE}

input_data <- read_csv("input_data/input_data.csv")

# only considering those inputs that lasted for more than a single frame =0.033s and shorter than 2.45s (time for instance to completely be solved) -> no guarantee to pursue goals but steer blindly
#input_data <- input_data[input_data$input_duration > 0.033, ]
input_data <- input_data[input_data$input_duration > 0.033 & input_data$input_duration < 2.45, ]
input_data <- input_data %>% filter(!is.na(input_duration))

input_data$input_noise <- as.factor(input_data$input_noise)
input_data <- within(input_data, input_noise <- relevel(input_noise, ref = "N"))

```

```{r input_duration_input, include=TRUE}

hist(input_data$input_duration, breaks=200)

fit.gamma <- fitdist(c(na.exclude(input_data$input_duration)), distr = "gamma", method = "mme")
summary(fit.gamma)
plot(fit.gamma)

```

```{r random_effects, include=TRUE}

inputDur_null <- glmer(input_duration ~ 1 + (1|ID), data=input_data, family = Gamma)
summary(inputDur_null)

```

```{r ICC_inputDur, include=TRUE}

ICC_inputDur <- 0.7201 / (0.7201 + 0.7758)
ICC_inputDur

```

```{r inputDur, include=TRUE}

inputDur <- glmer(input_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=input_data, family = Gamma)

inputDur.slope <- glmer(input_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=input_data, family = Gamma)

inputDur.slope2 <- glmer(input_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=input_data, family = Gamma)

inputDur.slope3 <- glmer(input_duration ~ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=input_data, family = Gamma)  # does not converge 

anova(inputDur, inputDur.slope, inputDur.slope2, inputDur.slope3)

# models significantly different
anova(inputDur, inputDur.slope2)

summary(inputDur.slope2)

```

```{r plot_fixation_duration_explore, echo=FALSE}

plot_inputDur <- plot_model(inputDur.slope2, 
           type = "pred", 
           terms = c("N_visible_obstacles", "N_visible_drift_tiles", "input_noise"),
           title = "",
           colors = "Set2",
           legend.title = TeX("$N_{ drift_{ visible}}$")
           )

# additional layout stuff
plot_inputDur <- plot_inputDur + 
  xlab(TeX("$N_{ obstacles_{ visible}}$")) +
  xlim(0, 15) +
  ylab("Input duration (s)")

plot_inputDur
```

# Responses to SoC questionnaire after each trial

```{r soc, include=TRUE}

soc_data <- read_csv("soc_data/soc_data.csv")

soc_data <- soc_data %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
soc_data$level_difficulty <- as.factor(soc_data$level_difficulty)
soc_data <- within(soc_data, level_difficulty <- relevel(level_difficulty, ref = "easy"))

soc_data$input_noise <- as.factor(soc_data$input_noise)
soc_data <- within(soc_data, input_noise <- relevel(input_noise, ref = "N"))

table(soc_data$SoC)

```

```{r soc_fit, include=TRUE}

hist(soc_data$SoC, breaks=7)

fit.gauss <- fitdist(c(na.exclude(soc_data$SoC)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

# AIC_gamma: 5600.357
# AIC_norm: 5124.438
```

```{r random_effects, include=TRUE}

soc_null <- lmer(SoC ~ 1 + (1|ID), data=soc_data, REML = FALSE)
summary(soc_null)

```

```{r ICC_soc, include=TRUE}

ICC_soc <- 0.3471 / (0.3471 + 2.5151)
ICC_soc

```

```{r soc, include=TRUE}

soc_model <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1|ID),  data = soc_data, REML = FALSE)

soc_model.slope <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+done|ID),  data = soc_data, REML = FALSE)

soc_model.slope2 <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+drift*level_difficulty|ID),  data = soc_data, REML = FALSE)  # does not converge

soc_model.slope3 <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+input_noise|ID),  data = soc_data, REML = FALSE)  # isSingular...

anova(soc_model, soc_model.slope, soc_model.slope2, soc_model.slope3)

summary(soc_model.slope)

```

Discussing the results: ...
