---
title: "linear_modeling_moonlander_eye_data"
author: "Nils Wendel Heinrich"
date: '2023-02-17'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(latex2exp)
library(ggplot2)
library(fitdistrplus)
library(simr)
library(arrow)

set.seed(36)
N_iterations <- 10000
```


```{r eye_data, include=FALSE}

setwd('/Users/heinrich/Projects/eye_movement_analysis_moonlander/')

data_fixations <- read_csv("eye_data/experimental_eye_data_fixations.csv")
data_saccades <- read_csv("eye_data/experimental_eye_data_saccades.csv")

```
We have data from 27 participants. Exclusion took place and will have to be specified here...

```{r edit_data, include=FALSE}

# level difficulty
# Most likely we won't use level difficulty as fixed effect. But we will nevertheless relevel it.
## fixations
data_fixations <- data_fixations %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
data_fixations$level_difficulty <- as.factor(data_fixations$level_difficulty)
data_fixations <- within(data_fixations, level_difficulty <- relevel(level_difficulty, ref = "easy"))
## saccades
data_saccades <- data_saccades %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
data_saccades$level_difficulty <- as.factor(data_saccades$level_difficulty)
data_saccades <- within(data_saccades, level_difficulty <- relevel(level_difficulty, ref = "easy"))

# input noise
## fixations
data_fixations$input_noise <- factor(data_fixations$input_noise, levels = c("N", "W", "S"))
#data_fixations$input_noise <- as.factor(data_fixations$input_noise)
#data_fixations <- within(data_fixations, input_noise <- relevel(input_noise, ref = "N"))
## saccades
data_saccades$input_noise <- factor(data_saccades$input_noise, levels = c("N", "W", "S"))
#data_saccades$input_noise <- as.factor(data_saccades$input_noise)
#data_saccades <- within(data_saccades, input_noise <- relevel(input_noise, ref = "N"))

```

# Effects coding

We are especially interested in exlporing effects of *input noise* in our predicted variables. Input noise means the internal loss of control. We will put it against *drift*, the external loss of control as well as account for the number of obstacles on the screen which might mainly drive changes in the predicted variable.

We want to apply our own effects coding for the variables of level_difficulty and input_noise.
```{r effects_coding, include=TRUE}

# both variables have 3 levels. Therefore we can go for a single matrix. We want to test level 2 against level 1, then level 3 against level 1, and last level 3 against level 2.
#effects_coding <- contr.treatment(3)

#effects_coding <- cbind( c(0,1,0), c(0,0,1), c(0,-1,1))
#contrasts(data_fixations$input_noise) = effects_coding

# this changes the base to level 2
#data_fixations$input_noise <- as.factor(data_fixations$input_noise)
#contrasts(data_fixations$input_noise) = contr.treatment(3, base = 2)

```

# Predicting fixations

Fixations may have served various functions. We will try to account for it and differentiate between resting and exploring fixations. 
```{r splitting_fixations, include=TRUE}

# only considering fixations with duration > 0:
data_fixations <- data_fixations[data_fixations$fixation_duration > 0, ]
# The eye tracker might have flagged single frames as fixations. These would have lasted no time and we will simply filter them out.

# distinguishing exploring and resting fixations
#x_sp = 972
#y_sp = 288

# 4° is half of the span of the parafovea. We will use that as the threshold (keeping the spaceship clearly visible vs. not)
#threshold = 4

#data_fixations$distance_to_spaceship <- sqrt((data_fixations$converging_eye_x_adjusted - x_sp)**2 + (data_fixations$converging_eye_y_adjusted - y_sp)**2)

exploring_fixations <- data_fixations[data_fixations$exploring_fixation == 1, ]
resting_fixations <- data_fixations[data_fixations$exploring_fixation == 0, ]

```

## Resting fixations - Fixation duration (s)

```{r check_fixDur_rest_fit, include=TRUE}

hist(resting_fixations$fixation_duration, breaks=200)

resting_fixations$fixation_duration.log <- log(resting_fixations$fixation_duration)

fit.gauss <- fitdist(resting_fixations$fixation_duration.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

fixDur_rest_null <- lmer(fixation_duration.log ~ 1 + (1|ID), data=resting_fixations, REML = FALSE)
summary(fixDur_rest_null)

```

```{r ICC_fixDur_rest, include=TRUE}

ICC_fixDur_rest <- 0.9733 / (0.9733 + 1.7213)
ICC_fixDur_rest

```

```{r fixation_duration_rest, include=TRUE}

#pred_fixDur_rest_complex <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1 + N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=resting_fixations, REML = FALSE)
# singularity = overparameterized

#pred_fixDur_rest_complex_ni <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1 + N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=resting_fixations, REML = FALSE)
# does not converge

pred_fixDur_rest <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=resting_fixations, REML = FALSE)

#pred_fixDur_rest.slope <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=resting_fixations, REML = FALSE) 
# model does not converge

pred_fixDur_rest.slope2 <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=resting_fixations, REML = FALSE)

pred_fixDur_rest.slope3 <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=resting_fixations, REML = FALSE)

#pred_fixDur_rest.slope4 <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles+input_noise|ID), data=resting_fixations, REML = FALSE)
# does not converge

#anova(pred_fixDur_rest, pred_fixDur_rest.slope, pred_fixDur_rest.slope2)

anova(pred_fixDur_rest, pred_fixDur_rest.slope2, pred_fixDur_rest.slope3)

# selecting pred_fixDur_rest.slope2 on basis of smallest BIC
summary(pred_fixDur_rest.slope2)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_fixDur_rest.slope2)

# bootstrap
confint(pred_fixDur_rest.slope2, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

## Resting fixations - Fixation location (distance to spaceship in visual degrees)

```{r check_distSpaceship_rest_fit, include=TRUE}

hist(resting_fixations$distance_to_spaceship, breaks=200)

fit.gauss <- fitdist(c(na.exclude(resting_fixations$distance_to_spaceship)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distSpaceship_rest_null <- lmer(distance_to_spaceship ~ 1 + (1|ID), data=resting_fixations, REML=FALSE)
summary(distSpaceship_rest_null)

```

```{r ICC_distSpaceship_rest, include=TRUE}

ICC_distSpaceship_rest <- 0.08396 / (0.08396 + 1.01570)
ICC_distSpaceship_rest

```

```{r distance_to_spaceship_rest, include=TRUE}

#pred_distSpaceship_rest_complex <- lmer(distance_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1 + N_visible_obstacles * N_visible_drift_tiles + input_noise |ID), data=resting_fixations, REML = FALSE)
# does not converge

#pred_distSpaceship_rest_complex_ni <- lmer(distance_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1 + N_visible_obstacles + N_visible_drift_tiles + input_noise |ID), data=resting_fixations, REML = FALSE)
# does not converge

#pred_distSpaceship_rest_complex_zc <- lmer(distance_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1 + N_visible_obstacles + N_visible_drift_tiles + input_noise||ID), data=resting_fixations, REML = FALSE)
# does not converge

pred_distSpaceship_rest <- lmer(distance_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=resting_fixations, REML = FALSE)

#pred_distSpaceship_rest.slope <- lmer(distance_to_spaceship ~ 1+ N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=resting_fixations, REML = FALSE)  # does not converge

pred_distSpaceship_rest.slope2 <- lmer(distance_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=resting_fixations, REML = FALSE)

#pred_distSpaceship_rest.slope3 <- lmer(distance_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=resting_fixations, REML = FALSE)
# does not converge

pred_distSpaceship_rest.slope4 <- lmer(distance_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles+input_noise|ID), data=resting_fixations, REML = FALSE)

anova(pred_distSpaceship_rest, pred_distSpaceship_rest.slope2, pred_distSpaceship_rest.slope4)

# selecting pred_distSpaceship_rest.slope4 due to smallest BIC
summary(pred_distSpaceship_rest.slope4)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distSpaceship_rest.slope4)

# bootstrap
confint(pred_distSpaceship_rest.slope4, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

## Exploring fixations - Fixation duration (s)

```{r check_fixDur_explore_fit, include=TRUE}

hist(exploring_fixations$fixation_duration, breaks=200)

exploring_fixations$fixation_duration.log <- log(exploring_fixations$fixation_duration)

fit.gauss <- fitdist(exploring_fixations$fixation_duration.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

fixDur_explore_null <- lmer(fixation_duration ~ 1 + (1|ID), data=exploring_fixations, REML = FALSE)
summary(fixDur_explore_null)

```

```{r ICC_fixDur_explore, include=TRUE}

ICC_fixDur_explore <- 0.02030 / (0.02030 + 0.09249)
ICC_fixDur_explore

```

```{r fixation_duration_explore, include=TRUE}

#pred_fixDur_explore_complex <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=exploring_fixations, REML = FALSE)
# does not converge

#pred_fixDur_explore_complex_ni <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=exploring_fixations, REML = FALSE)
# does not converge

pred_fixDur_explore <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=exploring_fixations, REML = FALSE)

pred_fixDur_explore.slope <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=exploring_fixations, REML = FALSE)

pred_fixDur_explore.slope2 <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=exploring_fixations, REML = FALSE)

#pred_fixDur_explore.slope3 <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=exploring_fixations, REML = FALSE) 
# model fails to converge

#pred_fixDur_explore.slope4 <- lmer(fixation_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles+N_visible_drift_tiles|ID), data=exploring_fixations, REML = FALSE)
# does not converge

anova(pred_fixDur_explore, pred_fixDur_explore.slope, pred_fixDur_explore.slope2)

# selecting pred_fixDur_explore.slope on basis of smallest BIC
summary(pred_fixDur_explore.slope)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_fixDur_explore.slope)

# bootstrap
confint(pred_fixDur_explore.slope, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

## Exploring fixations - Fixation location (distance to spaceship in visual degrees)

```{r check_distSpaceship_explore_fit, include=TRUE}

hist(exploring_fixations$distance_to_spaceship, breaks=200)

exploring_fixations$distance_to_spaceship.log <- log(exploring_fixations$distance_to_spaceship)

fit.gauss <- fitdist(c(na.exclude(exploring_fixations$distance_to_spaceship.log)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

#lnorm: AIC:  374016.8   BIC:  374034.8
#gamma: AIC:  400144.8   BIC:  400162.8
#exp: AIC:  407156.6   BIC:  407165.6

#log-transformed with norm: AIC:  103144.4   BIC:  103162.4

```

```{r random_effects, include=TRUE}

distSpaceship_explore_null <- lmer(distance_to_spaceship.log ~ 1 + (1|ID), data=exploring_fixations, REML = FALSE)
summary(distSpaceship_explore_null)

```

```{r ICC_distSpaceship_explore, include=TRUE}

ICC_distSpaceship_explore <- 0.05163 / (0.05163 + 0.24074)
ICC_distSpaceship_explore

```

```{r distance_to_spaceship_explore, include=TRUE}

#pred_distSpaceship_explore_complex <- lmer(distance_to_spaceship.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+ N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=exploring_fixations, REML = FALSE)
# does not converge

#pred_distSpaceship_explore_complex_ni <- lmer(distance_to_spaceship.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+ N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=exploring_fixations, REML = FALSE)
# does not converge

pred_distSpaceship_explore <- lmer(distance_to_spaceship.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=exploring_fixations, REML = FALSE)

#pred_distSpaceship_explore.slope <- lmer(distance_to_spaceship.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=exploring_fixations, REML = FALSE)
#model fails to converge

pred_distSpaceship_explore.slope2 <- lmer(distance_to_spaceship.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=exploring_fixations, REML = FALSE)

#pred_distSpaceship_explore.slope3 <- lmer(distance_to_spaceship.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=exploring_fixations, REML = FALSE)
# does not converge

anova(pred_distSpaceship_explore, pred_distSpaceship_explore.slope2)

# selecting pred_distSpaceship_explore.slope2 by referring to BIC
summary(pred_distSpaceship_explore.slope2)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distSpaceship_explore.slope2)

# bootstrap
confint(pred_distSpaceship_explore.slope2, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

# Predicting saccades

Saccades may have served various functions. We will try to account for it and differentiate between progressive and regressive saccades. 
```{r splitting_saccades, include=TRUE}

data_saccades <- data_saccades[data_saccades$saccade_amplitude > 0 & data_saccades$saccade_amplitude < 27, ]
# filter saccades that have no saccade amplitude
data_saccades <- data_saccades[!is.na(data_saccades$saccade_amplitude),]

progressive_saccades <- data_saccades[data_saccades$saccade_direction_y < 0, ]
regressive_saccades <- data_saccades[data_saccades$saccade_direction_y > 0, ]

```

## Progressive saccades - Saccade amplitude (in visual degrees)

```{r check_saccAmp_progress_fit, include=TRUE}

hist(progressive_saccades$saccade_amplitude, breaks=200)

progressive_saccades$saccade_amplitude.log <- log(progressive_saccades$saccade_amplitude)

fit.gauss <- fitdist(progressive_saccades$saccade_amplitude.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

#Gamma: AIC:  137235.3   BIC:  137251.8 
#log-transformed with norm: AIC:  101111.7   BIC:  101128.1 

```

```{r random_effects, include=TRUE}

saccAmp_progress_null <- lmer(saccade_amplitude.log ~ 1 + (1|ID), data=progressive_saccades, REML = FALSE)
summary(saccAmp_progress_null)

```

```{r ICC_saccAmp_progress, include=TRUE}

ICC_saccAmp_progress <- 0.213 / (0.213 + 2.101)
ICC_saccAmp_progress

```

```{r saccAmp_progress, include=TRUE}

#pred_saccAmp_progress_complex <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=progressive_saccades, REML = FALSE)
# singularity => overparameterized

#pred_saccAmp_progress_complex_ni <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=progressive_saccades, REML = FALSE)
# does not converge

pred_saccAmp_progress <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=progressive_saccades, REML = FALSE)

#pred_saccAmp_progress.slope <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=progressive_saccades, REML = FALSE) 
# model fails to converge

pred_saccAmp_progress.slope2 <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=progressive_saccades, REML = FALSE)

#pred_saccAmp_progress.slope3 <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=progressive_saccades, REML = FALSE)
# does not converge

anova(pred_saccAmp_progress, pred_saccAmp_progress.slope2)

# pred_saccAmp_progress.slope2 reached smaller BIC
summary(pred_saccAmp_progress.slope2)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_saccAmp_progress.slope2)

# bootstrap
confint(pred_saccAmp_progress.slope2, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

## Progressive saccades - distance to closest obstacle (in visual degrees)

```{r check_distObs_progress_fit, include=TRUE}

hist(progressive_saccades$saccLandSite_dist_to_closestObstacle, breaks=200)

hist(log(progressive_saccades$saccLandSite_dist_to_closestObstacle), breaks=200)

progressive_saccades$saccLandSite_dist_to_closestObstacle.log <- log(progressive_saccades$saccLandSite_dist_to_closestObstacle)

fit.gauss <- fitdist(c(na.exclude(progressive_saccades$saccLandSite_dist_to_closestObstacle.log)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distObs_explore_null <- lmer(saccLandSite_dist_to_closestObstacle.log ~ 1 + (1|ID), data=progressive_saccades, REML = FALSE)
summary(distObs_explore_null)

```

```{r ICC_distObs_progress, include=TRUE}

ICC_distObs_progress <- 0.1006 / (0.1006 + 0.7920)
ICC_distObs_progress

```

```{r distObs_progress, include=TRUE}

#pred_distObs_progress_complex <- lmer(saccLandSite_dist_to_closestObstacle.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=progressive_saccades, REML = FALSE)
# does not converge

#pred_distObs_progress_complex_ni <- lmer(saccLandSite_dist_to_closestObstacle.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=progressive_saccades, REML = FALSE)
# does not converge

pred_distObs_progress <- lmer(saccLandSite_dist_to_closestObstacle.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=progressive_saccades, REML = FALSE)

#pred_distObs_progress.slope <- lmer(saccLandSite_dist_to_closestObstacle.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=progressive_saccades, REML = FALSE)
# model fails to converge

pred_distObs_progress.slope2 <- lmer(saccLandSite_dist_to_closestObstacle.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=progressive_saccades, REML = FALSE)

pred_distObs_progress.slope3 <- lmer(saccLandSite_dist_to_closestObstacle.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=progressive_saccades, REML = FALSE)

anova(pred_distObs_progress, pred_distObs_progress.slope2, pred_distObs_progress.slope3)

# selecting model with smallest BIC: pred_distObs_progress.slope3
summary(pred_distObs_progress.slope3)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distObs_progress.slope3)

# bootstrap
confint(pred_distObs_progress.slope3, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

## Regressive saccades - Saccade amplitude

```{r check_saccAmp_regress_fit, include=TRUE}

hist(regressive_saccades$saccade_amplitude, breaks=200)

hist(log(regressive_saccades$saccade_amplitude), breaks=200)

regressive_saccades$saccade_amplitude.log <- log(regressive_saccades$saccade_amplitude)

fit.gauss <- fitdist(regressive_saccades$saccade_amplitude.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

saccAmp_regress_null <- lmer(saccade_amplitude ~ 1 + (1|ID), data=regressive_saccades, REML = FALSE)
summary(saccAmp_regress_null)

```

```{r ICC_saccAmp_regress, include=TRUE}

ICC_saccAmp_regress <- 0.6771 / (0.6771 + 15.3863)
ICC_saccAmp_regress

```

```{r saccAmp_regress, include=TRUE}

#pred_saccAmp_regress_complex <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+ N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=regressive_saccades, REML = FALSE)
# singularity => overparameterized

#pred_saccAmp_regress_complex_ni <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+ N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=regressive_saccades, REML = FALSE)
# singularity => overparameterized

# throwing out input noise
#pred_saccAmp_regress_complex_ni <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+ N_visible_obstacles + N_visible_drift_tiles|ID), data=regressive_saccades, REML = FALSE)
# does not converge

pred_saccAmp_regress <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=regressive_saccades, REML = FALSE)

pred_saccAmp_regress.slope <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=regressive_saccades, REML = FALSE)

pred_saccAmp_regress.slope2 <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=regressive_saccades, REML = FALSE)

pred_saccAmp_regress.slope3 <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=regressive_saccades, REML = FALSE)

#pred_saccAmp_regress.slope_obsin <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+ N_visible_obstacles + input_noise|ID), data=regressive_saccades, REML = FALSE)
# does not converge

pred_saccAmp_regress.slope_driftin <- lmer(saccade_amplitude.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+ N_visible_drift_tiles + input_noise|ID), data=regressive_saccades, REML = FALSE)

anova(pred_saccAmp_regress, pred_saccAmp_regress.slope, pred_saccAmp_regress.slope2, pred_saccAmp_regress.slope3, pred_saccAmp_regress.slope_driftin)

# selecting pred_saccAmp_regress.slope due to smallest BIC
summary(pred_saccAmp_regress.slope)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_saccAmp_regress.slope)

# bootstrap
confint(pred_saccAmp_regress.slope, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

## Regressive saccades - distance to spaceship

```{r check_distSpaceship_regress_fit, include=TRUE}

hist(regressive_saccades$saccLandSite_dist_to_spaceship, breaks=200)

fit.gauss <- fitdist(regressive_saccades$saccade_amplitude, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distSpaceship_regress_null <- lmer(saccLandSite_dist_to_spaceship ~ 1 + (1|ID), data=regressive_saccades, REML = FALSE)
summary(distSpaceship_regress_null)

```

```{r ICC_distSpaceship_regress, include=TRUE}

ICC_distSpaceship_regress <- 53.49 / (53.49 + 234.88)
ICC_distSpaceship_regress

```

```{r distSpaceship_regress, include=TRUE}

#pred_distSpaceship_regress_complex <- lmer(saccLandSite_dist_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=regressive_saccades, REML = FALSE)
# model does not converge

#pred_distSpaceship_regress_complex_ni <- lmer(saccLandSite_dist_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=regressive_saccades, REML = FALSE)
# does not converge

pred_distSpaceship_regress <- lmer(saccLandSite_dist_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=regressive_saccades, REML = FALSE)

#pred_distSpaceship_regress.slope <- lmer(saccLandSite_dist_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=regressive_saccades, REML = FALSE) # does not converge

pred_distSpaceship_regress.slope2 <- lmer(saccLandSite_dist_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=regressive_saccades, REML = FALSE)

pred_distSpaceship_regress.slope3 <- lmer(saccLandSite_dist_to_spaceship ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=regressive_saccades, REML = FALSE)

anova(pred_distSpaceship_regress, pred_distSpaceship_regress.slope2, pred_distSpaceship_regress.slope3)

# pred_distSpaceship_regress.slope3 reached smallest BIC
summary(pred_distSpaceship_regress.slope3)

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distSpaceship_regress.slope3)

# bootstrap
confint(pred_distSpaceship_regress.slope3, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

Discussing the results: ...

# Input data - input duration

```{r input_data, include=FALSE}

input_data <- read_csv("input_data/input_data.csv")

# only considering those inputs that lasted for more than a single frame =0.033s and shorter than 2.45s (time for instance to completely be solved) -> no guarantee to pursue goals but steer blindly
#input_data <- input_data[input_data$input_duration > 0.033, ]
input_data <- input_data[input_data$input_duration > 0.033 & input_data$input_duration < 2.45, ]
input_data <- input_data %>% filter(!is.na(input_duration))

input_data$input_noise <- as.factor(input_data$input_noise)
input_data <- within(input_data, input_noise <- relevel(input_noise, ref = "N"))

```

```{r input_duration_input, include=TRUE}

hist(input_data$input_duration, breaks=200)

hist(log(input_data$input_duration), breaks=200)

input_data$input_duration.log <- log(input_data$input_duration)

fit.gauss <- fitdist(c(na.exclude(input_data$input_duration.log)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

inputDur_null <- lmer(input_duration.log ~ 1 + (1|ID), data=input_data, REML = FALSE)
summary(inputDur_null)

```

```{r ICC_inputDur, include=TRUE}

ICC_inputDur <- 0.05978 / (0.05978 + 0.47762)
ICC_inputDur

```

```{r inputDur, include=TRUE}

#inputDur_complex <- lmer(input_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles * N_visible_drift_tiles + input_noise|ID), data=input_data, REML = FALSE)
# model does not converge

#inputDur_complex_ni <- lmer(input_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles + N_visible_drift_tiles + input_noise|ID), data=input_data, REML = FALSE)
# does not converge

inputDur <- lmer(input_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1|ID), data=input_data, REML = FALSE)

#inputDur.slope <- lmer(input_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_obstacles|ID), data=input_data, REML = FALSE) # model fails to converge

inputDur.slope2 <- lmer(input_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles|ID), data=input_data, REML = FALSE)

inputDur.slope3 <- lmer(input_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+input_noise|ID), data=input_data, REML = FALSE)

inputDur.slope_dirftin <- lmer(input_duration.log ~ 1 + N_visible_obstacles * N_visible_drift_tiles + input_noise + (1+N_visible_drift_tiles+input_noise|ID), data=input_data, REML = FALSE)

anova(inputDur, inputDur.slope2, inputDur.slope3, inputDur.slope_dirftin)

# inputDur.slope2 reached smallest BIC
summary(inputDur.slope2)

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(inputDur.slope2)

# bootstrap
confint(inputDur.slope2, nsim=N_iterations, parm=c("N_visible_obstacles", "N_visible_drift_tiles", "input_noiseW", "input_noiseS", "N_visible_obstacles:N_visible_drift_tiles"), method="boot")

```

# Responses to SoC questionnaire after each trial

```{r soc, include=TRUE}

soc_data <- read_csv("soc_data/soc_data.csv")

soc_data <- soc_data %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
soc_data$level_difficulty <- as.factor(soc_data$level_difficulty)
soc_data <- within(soc_data, level_difficulty <- relevel(level_difficulty, ref = "easy"))

soc_data$input_noise <- as.factor(soc_data$input_noise)
soc_data <- within(soc_data, input_noise <- relevel(input_noise, ref = "N"))

table(soc_data$SoC)

```

```{r soc_fit, include=TRUE}

hist(soc_data$SoC, breaks=7)

fit.gauss <- fitdist(c(na.exclude(soc_data$SoC)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

# AIC_gamma: 5600.357
# AIC_norm: 5124.438
```

```{r random_effects, include=TRUE}

soc_null <- lmer(SoC ~ 1 + (1|ID), data=soc_data, REML = FALSE)
summary(soc_null)

```

```{r ICC_soc, include=TRUE}

ICC_soc <- 0.3287 / (0.3287 + 2.4395)
ICC_soc

```

```{r soc, include=TRUE}

soc_model <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1|ID),  data = soc_data, REML = FALSE)

soc_model.slope <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+done|ID),  data = soc_data, REML = FALSE)

#soc_model.slope_in <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+drift*level_difficulty|ID),  data = soc_data, REML = FALSE)  
# isSingular... therefore overparameterized

#soc_model.slope_ni <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+drift+level_difficulty|ID),  data = soc_data, REML = FALSE) 
# singularity => overparamterized

soc_model.slope2 <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+drift|ID),  data = soc_data, REML = FALSE)

#soc_model.slope3 <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+level_difficulty|ID),  data = soc_data, REML = FALSE)
# singularity => overparamterized

#soc_model.slope4 <- lmer(SoC ~ done + drift*level_difficulty + input_noise + (1+input_noise|ID),  data = soc_data, REML = FALSE)  
# isSingular... therefore overparameterized

anova(soc_model, soc_model.slope, soc_model.slope2)

# selecting soc_model.slope by referring to BIC
summary(soc_model.slope)

```

Discussing the results: ...

# Exploring effects in SoC response

```{r soc, include=TRUE}

soc_data <- read_csv("soc_data/soc_data.csv")

soc_data <- soc_data %>%
  mutate(level_difficulty = case_when(
    (level == 1 | level == 2) ~ "easy",
    (level == 3 | level == 4) ~ "medium",
    (level == 5 | level == 6) ~ "hard",
    ))
soc_data$level_difficulty <- as.factor(soc_data$level_difficulty)
soc_data <- within(soc_data, level_difficulty <- relevel(level_difficulty, ref = "easy"))

soc_data$input_noise <- as.factor(soc_data$input_noise)
soc_data <- within(soc_data, input_noise <- relevel(input_noise, ref = "N"))

#table(soc_data$SoC)

write_feather(soc_data, "soc_data/Experiment1_SoCData.arrow")

```

```{r soc, include=TRUE}

#soc_model_complex <- lmer(SoC ~ done + drift*level_difficulty + input_noise + N_prior_crashs + (1+done+ drift*level_difficulty + input_noise + N_prior_crashs|ID),  data = soc_data, REML = FALSE)
# overparameterized

# deleting interaction term within random effects structue
#soc_model_complex_ni <- lmer(SoC ~ done + drift*level_difficulty + input_noise + N_prior_crashs + (1+done+ drift + level_difficulty + input_noise + N_prior_crashs|ID),  data = soc_data, REML = FALSE)
# overparameterized

# deleting correlation between random effects
#soc_model_complex_nc <- lmer(SoC ~ done + drift*level_difficulty + input_noise + N_prior_crashs + (1 + done + drift + level_difficulty + input_noise + N_prior_crashs||ID),  data = soc_data, REML = FALSE)
# overparameterized

soc_model <- lmer(SoC ~ done + drift*level_difficulty + input_noise + N_prior_crashs + (1+done|ID),  data = soc_data, REML = FALSE)
#summary(soc_model) # N_prior_crashs is significant but positively correlated 

#soc_model_in <- lmer(SoC ~ done + drift*level_difficulty + input_noise + trials_since_last_crash + (1+drift*level_difficulty|ID),  data = soc_data, REML = FALSE)
# does not converge

#soc_model_driftdiff <- lmer(SoC ~ done + drift*level_difficulty + input_noise + trials_since_last_crash + (1+drift+level_difficulty|ID),  data = soc_data, REML = FALSE)
# singular => overparameterized

#soc_model_indone <- lmer(SoC ~ done + drift*level_difficulty + input_noise + trials_since_last_crash + (1+done+input_noise|ID),  data = soc_data, REML = FALSE)
# singular => overparameterized

soc_model <- lmer(SoC ~ done + drift*level_difficulty + input_noise + crashed_in_last_trial + (1+done|ID),  data = soc_data, REML = FALSE)
summary(soc_model) # trials_since_last_crash not significant

############# significant effect - keeping random effects structure

soc_model.slope <- lmer(SoC ~ done + drift*level_difficulty + input_noise + N_consecutive_crash_success + (1+done|ID),  data = soc_data, REML = FALSE)

#soc_model.slope2 <- lmer(SoC ~ done + drift*level_difficulty + input_noise + N_consecutive_crash_success + (1+N_consecutive_crash_success|ID),  data = soc_data, REML = FALSE)
# does not converge

```


# Preparing data of 2nd experiment:

```{r soc, include=TRUE}

fixData_exp2 <- read_csv("eye_data/experimental_eye_data_fixations_experiment2.csv")
saccData_exp2 <- read_csv("eye_data/experimental_eye_data_saccades_experiment2.csv")

# Fixations
# only considering fixations with duration > 0:
fixData_exp2 <- fixData_exp2[fixData_exp2$fixation_duration > 0, ]
# The eye tracker might have flagged single frames as fixations. These would have lasted no time and we will simply filter them out.

# distinguishing exploring and resting fixations
#x_sp = 972
#y_sp = 288

# 4° is half of the span of the parafovea. We will use that as the threshold (keeping the spaceship clearly visible vs. not)
#threshold = 4

#data_fixations$distance_to_spaceship <- sqrt((data_fixations$converging_eye_x_adjusted - x_sp)**2 + (data_fixations$converging_eye_y_adjusted - y_sp)**2)

exploring_fixations_exp2 <- fixData_exp2[fixData_exp2$exploring_fixation == 1, ]
resting_fixations_exp2 <- fixData_exp2[fixData_exp2$exploring_fixation == 0, ]

###############################################################################
# Saccades
saccData_exp2 <- saccData_exp2[saccData_exp2$saccade_amplitude > 0 & saccData_exp2$saccade_amplitude < 27, ]
# filter saccades that have no saccade amplitude
saccData_exp2 <- saccData_exp2[!is.na(saccData_exp2$saccade_amplitude),]

progressive_saccades_exp2 <- saccData_exp2[saccData_exp2$saccade_direction_y < 0, ]
regressive_saccades_exp2 <- saccData_exp2[saccData_exp2$saccade_direction_y > 0, ]


# Writing to arrow
write_feather(exploring_fixations_exp2, "Experiment2_ExploringFixations.arrow")
write_feather(resting_fixations_exp2, "Experiment2_RestingFixations.arrow")

write_feather(progressive_saccades_exp2, "Experiment2_ProgressiveSaccades.arrow")
write_feather(regressive_saccades_exp2, "Experiment2_RegressiveSaccades.arrow")
```








