---
title: "situational_analysis_linear_models"
author: "Nils Wendel Heinrich"
date: '2023-08-16'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(sjPlot)
library(ggplot2)
library(fitdistrplus)

set.seed(36)
N_iterations <- 10000
```

```{r data, include=FALSE}

setwd('/Users/heinrich/Projects/eye_movement_analysis_moonlander/')

data_fixations <- read_csv("eye_data/situational_analysis_data/situational_fixations_data.csv")
data_saccades <- read_csv("eye_data/situational_analysis_data/situational_saccades_data.csv")

```

# Predicting fixations

Variables of interest are:
- *fixation_duration*
- *distSpaceship_fix_rest*
- *distClosestObstacle_fix_explore*
whereby we have to differentiate between **resting** and **exploring** fixations.

We will start by subsetting the data into two different data sets. For this we will filter out fixations of 0 duration. These may occur due to the eye tracker flagging only a single frame as fixation.
```{r rest_vs_explore, include=TRUE}

# only considering fixations with duration > 0:
data_fixations <- data_fixations[data_fixations$fixation_duration > 0, ]

exploring_fixations <- data_fixations[data_fixations$exploring_fixation == TRUE, ]
resting_fixations <- data_fixations[data_fixations$exploring_fixation == FALSE, ]

```

## Resting fixations - Fixation duration

```{r check_fixDur_rest_fit, include=TRUE}

hist(resting_fixations$fixation_duration, breaks=200)

hist(log(resting_fixations$fixation_duration), breaks=200)

resting_fixations$fixation_duration.log <- log(resting_fixations$fixation_duration)

fit.gauss <- fitdist(resting_fixations$fixation_duration.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

#Gamma: AIC:  -589.6791   BIC:  -579.7603
#log-transformed normal: AIC:  4363.395   BIC:  4373.313

```

```{r random_effects, include=TRUE}

fixDur_rest_null <- lmer(fixation_duration.log ~ 1 + (1|code), data=resting_fixations, REML = FALSE)

summary(fixDur_rest_null)

```

```{r ICC_fixDur_rest, include=TRUE}

ICC_fixDur_rest <- 1.251 / (1.251 + 1.662)

ICC_fixDur_rest

```

```{r fixation_duration, include=TRUE}

# log normal
pred_fixDur_rest <- lmer(fixation_duration.log ~ success + (1|code), data=resting_fixations, REML = FALSE)

pred_fixDur_rest.slope <- lmer(fixation_duration.log ~ success + (1+success|code), data=resting_fixations, REML = FALSE)

anova(pred_fixDur_rest, pred_fixDur_rest.slope)

summary(pred_fixDur_rest.slope)

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_fixDur_rest.slope)

# bootstrap
confint(pred_fixDur_rest.slope, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```

## Resting fixations - Distance to spaceship

We will assume a normal distribution for this predicted variable.
```{r check_distSpaceship_fit, include=TRUE}

hist(resting_fixations$distSpaceship_fix_rest, breaks=200)

fit.gauss <- fitdist(resting_fixations$distSpaceship_fix_rest, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distSpaceship_null <- lmer(distSpaceship_fix_rest ~ 1 + (1|code), data=resting_fixations, REML=FALSE)
summary(distSpaceship_null)

```

```{r ICC_distSpaceship, include=TRUE}

ICC_distSpaceship <- 0.08552 / (0.08552 + 0.77587)
ICC_distSpaceship

```

```{r fixation_duration, include=TRUE}

pred_distSpaceship <- lmer(distSpaceship_fix_rest ~ success + (1|code), data=resting_fixations, REML=FALSE)

pred_distSpaceship.slope <- lmer(distSpaceship_fix_rest ~ success + (1+success|code), data=resting_fixations, REML=FALSE) # overparameterized

summary(pred_distSpaceship)

```

Discussing the results: We find a significant fixed effect for success on the distance to the spaceship in resting fixations. This means that in successful runs, participants allocated their resting fixations further away from the spaceship compared to crash runs.

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distSpaceship)

# bootstrap
confint(pred_distSpaceship, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```

## Exploring fixations - Fixation duration

```{r check_distSpaceship_fit, include=TRUE}

hist(exploring_fixations$fixation_duration, breaks=200)

hist(log(exploring_fixations$fixation_duration), breaks=200)

exploring_fixations$fixation_duration.log <- log(exploring_fixations$fixation_duration)

fit.gauss <- fitdist(exploring_fixations$fixation_duration.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

fixDur_explore_null <- lmer(fixation_duration.log ~ 1 + (1|code), data=exploring_fixations, REML = FALSE)

summary(fixDur_explore_null)

```

```{r ICC_fixDur_explore, include=TRUE}

ICC_fixDur_explore <- 1.069 / (1.069 + 1.886)

ICC_fixDur_explore

```

```{r fixation_duration, include=TRUE}

#pred_fixDur_explore <- glmer(fixation_duration ~ success + (1|code), data=exploring_fixations, family = Gamma)

pred_fixDur_explore <- lmer(fixation_duration.log ~ success + (1|code), data=exploring_fixations, REML = FALSE)

#pred_fixDur_explore.slope <- glmer(fixation_duration ~ success + (1+success|code), data=exploring_fixations, family = Gamma)

pred_fixDur_explore.slope <- lmer(fixation_duration.log ~ success + (1+success|code), data=exploring_fixations, REML = FALSE)

anova(pred_fixDur_explore, pred_fixDur_explore.slope)

summary(pred_fixDur_explore.slope)

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_fixDur_explore.slope)

# bootstrap
confint(pred_fixDur_explore.slope, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```

## Exploring fixations - Distance to closest obstacle

```{r check_distObstacle_fit, include=TRUE}

hist(exploring_fixations$distClosestObstacle_fix_explore, breaks=200)
# this looks like a truncated normal distribution

hist(log(exploring_fixations$distClosestObstacle_fix_explore), breaks=200)

exploring_fixations$distClosestObstacle_fix_explore.log <- log(exploring_fixations$distClosestObstacle_fix_explore)

fit.gauss <- fitdist(exploring_fixations$distClosestObstacle_fix_explore.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

#gauss: AIC:  23729.99   BIC:  23740.75
#gamma: AIC:  22488.64   BIC:  22499.4 


```

```{r random_effects, include=TRUE}

distObstacle_null <- lmer(distClosestObstacle_fix_explore.log ~ 1 + (1|code), data=exploring_fixations, REML = FALSE)

summary(distObstacle_null)

```

```{r ICC_distObstacle, include=TRUE}

ICC_distObstacle <- 0.08692 / (0.08692 + 0.14374)

ICC_distObstacle

```

```{r distClosestObstacle_fix_explore, include=TRUE}

pred_distObstacle <- lmer(distClosestObstacle_fix_explore.log ~ success + (1|code), data=exploring_fixations, REML = FALSE)

pred_distObstacle.slope <- lmer(distClosestObstacle_fix_explore.log ~ success + (1+success|code), data=exploring_fixations, REML = FALSE)

anova(pred_distObstacle, pred_distObstacle.slope)

summary(pred_distObstacle.slope)

```

Discussing the results: ...

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distObstacle.slope)

# bootstrap
confint(pred_distObstacle.slope, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```

# Predicting saccades

Variables of interest are:
- *saccade_amplitude*
- *distClosestObstacle_sacc_progress*
- *distClosestObstacle_sacc_regress*
whereby we have to differentiate between **progressive** and **regressive** saccades

We will start to subset the data into two different data sets. For this we will filter out saccades of 0 amplitude These may occur due to the eye tracker flagging only a single frame as saccade.
```{r progress_vs_regress, include=TRUE}

# only considering fixations with duration > 0:
data_saccades <- data_saccades[data_saccades$saccade_amplitude > 0, ]

progressive_saccades <- data_saccades[data_saccades$progressive_saccade == TRUE, ]
regressive_saccades <- data_saccades[data_saccades$progressive_saccade == FALSE, ]

```

## Progressive saccades - Saccade amplitude

```{r check_saccAmp_progress_fit, include=TRUE}

#hist(progressive_saccades$saccade_amplitude), breaks=200)

#fit.gamma <- fitdist(progressive_saccades$saccade_amplitude, distr = "gamma", method = "mme")
#summary(fit.gamma)
#plot(fit.gamma)

progressive_saccades$saccade_amplitude.log <- log(progressive_saccades$saccade_amplitude)

hist(log(progressive_saccades$saccade_amplitude), breaks=200)

fit.gauss <- fitdist(progressive_saccades$saccade_amplitude.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

saccAmp_progress_null <- lmer(saccade_amplitude.log ~ 1 + (1|code), data=progressive_saccades, REML = FALSE)

summary(saccAmp_progress_null)

```

```{r ICC_saccAmp_progress, include=TRUE}

ICC_saccAmp_progress <- 0.3076 / (0.3076 + 2.1804)

ICC_saccAmp_progress

```

```{r saccade_amplitude_progress, include=TRUE}

pred_saccAmp_progress <- lmer(saccade_amplitude.log ~ success + (1|code), data=progressive_saccades, REML = FALSE)

pred_saccAmp_progress.slope <- lmer(saccade_amplitude.log ~ success + (1+success|code), data=progressive_saccades, REML = FALSE)

anova(pred_saccAmp_progress, pred_saccAmp_progress.slope)

summary(pred_saccAmp_progress)

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_saccAmp_progress)

# bootstrap
confint(pred_saccAmp_progress, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```

## Progressive saccades - Distance to closest obstacle

```{r check_distObs_progress_fit, include=TRUE}

hist(progressive_saccades$distClosestObstacle_sacc_progress, breaks=200)

# maybe log transform distClosestObstacle columns?
progressive_saccades$distClosestObstacle_sacc_progress.log <- log(progressive_saccades$distClosestObstacle_sacc_progress)

# we cannot ensure that distances won't be =0 (not allowed in gamma). We therefore assume a normal distribution
fit.gauss <- fitdist(c(na.exclude(progressive_saccades$distClosestObstacle_sacc_progress.log)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distClosestObstacle_sacc_progress_null <- lmer(distClosestObstacle_sacc_progress.log ~ 1 + (1|code), data=progressive_saccades, REML=FALSE)
summary(distClosestObstacle_sacc_progress_null)

```

```{r ICC_distObs_progress, include=TRUE}

ICC_distObs_progress <- 0.2661 / (0.2661 + 0.8950)
ICC_distObs_progress

```

```{r distObstacle_progress, include=TRUE}

pred_distObstacle_progress <- lmer(distClosestObstacle_sacc_progress.log ~ success + (1|code), data=progressive_saccades, REML=FALSE)

pred_distObstacle_progress.slope <- lmer(distClosestObstacle_sacc_progress.log ~ success + (1+success|code), data=progressive_saccades, REML=FALSE)

anova(pred_distObstacle_progress, pred_distObstacle_progress.slope)

summary(pred_distObstacle_progress.slope)
# we loose the fixed effect for success when including the random slope effect for code on success, but the random slope model is the one with the better AIC & BIC

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distObstacle_progress.slope)

# bootstrap
confint(pred_distObstacle_progress.slope, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```

## Regressive saccades - Saccade amplitude

```{r check_saccAmp_regress_fit, include=TRUE}

hist(regressive_saccades$saccade_amplitude, breaks=200)

hist(log(regressive_saccades$saccade_amplitude), breaks=200)

regressive_saccades$saccade_amplitude.log <- log(regressive_saccades$saccade_amplitude)

fit.gauss <- fitdist(regressive_saccades$saccade_amplitude.log, distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

saccAmp_regress_null <- lmer(saccade_amplitude.log ~ 1 + (1|code), data=regressive_saccades, REML = FALSE)
summary(saccAmp_regress_null)

```

```{r ICC_saccAmp_regress, include=TRUE}

ICC_saccAmp_regress <- 0.2691 / (0.2691 + 1.7962)
ICC_saccAmp_regress

```

```{r saccade_amplitude_regress, include=TRUE}

pred_saccAmp_regress <- lmer(saccade_amplitude.log ~ success + (1|code), data=regressive_saccades, REML = FALSE)

pred_saccAmp_regress.slope <- lmer(saccade_amplitude.log ~ success + (1+success|code), data=regressive_saccades, REML = FALSE)

anova(pred_saccAmp_regress, pred_saccAmp_regress.slope)

summary(pred_saccAmp_regress)

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_saccAmp_regress)

# bootstrap
confint(pred_saccAmp_regress, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```

## Regressive saccades - Distance to closest obstacle

```{r check_distObs_fit, include=TRUE}

hist(regressive_saccades$distClosestObstacle_sacc_regress, breaks=200)

hist(log(regressive_saccades$distClosestObstacle_sacc_regress), breaks=200)

regressive_saccades$distClosestObstacle_sacc_regress.log <- log(regressive_saccades$distClosestObstacle_sacc_regress)

# we cannot ensure that distances won't be =0. We therefore assume a normal distribution
fit.gauss <- fitdist(c(na.exclude(regressive_saccades$distClosestObstacle_sacc_regress.log)), distr = "norm", method = "mme")
summary(fit.gauss)
plot(fit.gauss)

```

```{r random_effects, include=TRUE}

distClosestObstacle_sacc_regress_null <- lmer(distClosestObstacle_sacc_regress.log ~ 1 + (1|code), data=regressive_saccades, REML=FALSE)
summary(distClosestObstacle_sacc_regress_null)

```

```{r ICC_distObs_regress, include=TRUE}

ICC_distObs_regress <- 0.7747 / (0.7747 + 1.0511)
ICC_distObs_regress

```

```{r distObstacle_regress, include=TRUE}

pred_distObstacle_regress <- lmer(distClosestObstacle_sacc_regress.log ~ success + (1|code), data=regressive_saccades, REML=FALSE)

pred_distObstacle_regress.slope <- lmer(distClosestObstacle_sacc_regress.log ~ success + (1+success|code), data=regressive_saccades, REML=FALSE)

anova(pred_distObstacle_regress, pred_distObstacle_regress.slope)

# again, we loose the fixed effect for success when including the random slope effect for code on success, but the random slope model is the one with the better AIC & BIC
summary(pred_distObstacle_regress.slope)

```

```{r eval=TRUE, cache=TRUE}
# fixed effects
fixef(pred_distObstacle_regress.slope)

# bootstrap
confint(pred_distObstacle_regress.slope, nsim=N_iterations, parm=c("successTRUE"), method="boot")

```
